---
title: "Coffee"
author: "Hawra Nawrozzadeh"
date: "03/02/2022"
output: html_document
---

```{r}
coffee <- read.csv("data/Coffee-modified.csv", stringsAsFactors = T)
```

```{r}
str(coffee)
summary(coffee)
colnames(coffee)
```

- Maybe get rid of the followings:
  - Owner
  -hello
  - Farm Name
  - Lot Number
  - Mill 
  - ICO Number
  - Company
  - Region
  -hello
  - Producer
  - In country Partner
  - Harvest Year
  - Owner 1
  - Expiration
  - Certification Body
  - certification address 
  - Certification number
  - Unit of measurements
  
-  Columns to keep:
 - ID
 - Species
 - Orgin of Country
 - Altitude
 - No. of bags
 - Bag weight
 - Grading Date
 - Variety
 - Processing method
 - Aroma 
 - Flavor
 - Aftertaste
 - Acidity
 - Body
 - Balance
 - Uniformity
 - Clean.up
 - Sweetness
 - Cupper.Point
 - Total.Cup.point
 - Moisture
 - Quakers
 - colors
 - the last 3 
```{r}
View(table(coffee$Owner.1))
View(table(coffee$Country.of.Origin))
table(coffee$Variety)
table(coffee$Category.One.Defects)
table(coffee$Category.Two.Defects)


#coffee[ ,c("Species", "Altitude", "Number.of.Bags", "Bag.Weight", "Processing.Method", "Aroma", "Flavor", "Aftertaste", "Acidity", "Body", "Balance", "Uniformity", "Clean.Up", "Sweetness")]

coffee_new <- coffee[ ,c(1, 2, 4, 10, 13, 14, 17, 19:32, 34, 42:44)]
```

Data Preparation and Cleaning
  - Feature and Selection Process
  - `as.factor()`
  - Make sure the decimal columns are in the right data type
  - In general, checking the structure and check that they are all in the same data type
  - Maybe use validate package
  
  - Dealing with Outliers
  - Missing Data --> Imputation
  - Remove duplicates

```{r}
summary()
```

```{r}
coffee_1 <- read.csv("data/coffee.csv")
```


