---
title: "CS5811-Data Distributed Analysis-Data Analysis on Arabica Coffee Beans"
author: "Hawra Nawrozzadeh, Joao Pedro Azevedo, and Roberta Sammarco"
date: "10/02/2022"
output: html_document
version: 1.0
editor_options: 
  chunk_output_type: inline
---

# 1. Introduction

- data is on Arabic coffee
- show metadata
- Research Question: ***"Given the quality measures of the Arabica coffee beans, can we predict where these coffees are originated from?"***

# 2. Loading the Libraries

```{r}
# For data cleaning and preparation
library(dplyr)
library(tidyverse)
library(validate)
# For Data visualisation
library(ggplot2)
```

# 3. Data Cleaning and Preparation

## 3.1 Loading the data 

```{r}
coffee <- read.csv("Coffee-modified.csv")
```


## 3.2 Data Qaulity Check (or something like this)

Let's look into the structure of the data set
```{r}
str(coffee)
summary(coffee)
```

-From the `summary()` and `str()` function, it seems R has processed the dataset with  35 attributes categorical variables and the rest of the 9 attributes to be numerical continuous.
- However, this seems to be incorrect format because the in correct datatype..... e.g. ID and bla bla should be integer instead ...
- Additionally, we can already identify errors such as missing data and possible outliers (from looking at the ranges)...

- Lets further look into the different levels by creating an object that converts the strings into a factor. By doing this, we can also see other details that we will help us to decide which attributes are redundant or irrelavent for our research question.

- The `summary()` and `str()` function shows that the dataset contains

- 35 categorical/character variables
- 9 continuous variables
- incorrect data types
- cant get much from `summary()` function--> use `stringAsFactor = T` to get some more insight 

```{r}
coffee_fac_lev <- read.csv("Coffee-modified.csv", stringsAsFactors = T)
str(coffee_fac_lev)
summary(coffee_fac_lev)
```


- There is alot of errors and redudant information
- can see the amount of missings informations, etc. 
(maybe this is where we use the validate package?)
- As a results, in order to answer our research, we would only need the following variables....

----
à	ID
à	Owner
à	Farm Name
à	Lot Number
à	Mill 
à	ICO Number
à	Company
à	Region
à	Producer
à	In country Partner
à	Harvest Year
à	Owner 1
à	Expiration
à	Certification Body
à	certification address 
à	Certification number
à	Unit of measurements
^^ this is needs to be fixed accordingly

----

```{r}
colnames(coffee)

## should get rid of (processing methods), color, and species as these are associated with the bean metadata and not the part of the quality of the coffee, plus they have alot of missing rows

coffee_new <- coffee %>%
  select(-c( ## removing data regarding the Farm 
            "Owner", "Farm.Name", "Lot.Number", "Mill", "ICO.Number",
            "Company","Region","Producer","In.Country.Partner","Harvest.Year","Owner.1","Variety", "Expiration",
            "Certification.Body","Certification.Address","Certification.Contact","unit_of_measurement",
            ## removing data regard the bean
            "Color", "Species", "Altitude", "altitude_low_meters", "altitude_high_meters", "altitude_mean_meters", "Grading.Date"
            )
         )
colnames(coffee_new)
str(coffee_new)
summary(coffee_new)
```



- Assessing Empty Rows
```{r}
summary(coffee_new)
colSums(is.na(coffee_new))
colSums(coffee_new == "")

#https://stackoverflow.com/questions/42721788/filter-empty-rows-from-a-dataframe-with-r
coffee_new_1 <- coffee_new[Reduce(`&`, lapply(coffee_new, 
                                              function(x) !(x==""))),]

colSums(is.na(coffee_new_1))
colSums(coffee_new_1 == "")
```

Investigating further to identify the errors and see from there what is the appropriate approach
```{r}
table(coffee_new_1$Country.of.Origin)
View(table(coffee_new_1$Number.of.Bags))
table(coffee_new_1$Bag.Weight) ##
table(coffee_new_1$Processing.Method)
table(coffee_new_1$Aroma) ##
table(coffee_new_1$Flavor) 
table(coffee_new_1$Aftertaste) 
table(coffee_new_1$Acidity) 
table(coffee_new_1$Body)
table(coffee_new_1$Balance)
table(coffee_new_1$Uniformity) ##
table(coffee_new_1$Clean.Cup) ##
table(coffee_new_1$Sweetness) ##
table(coffee_new_1$Cupper.Points) ##
table(coffee_new_1$Total.Cup.Points) ##
table(coffee_new_1$Moisture) 
table(coffee_new_1$Category.One.Defects)
table(coffee_new_1$Quakers) ##
table(coffee_new_1$Category.Two.Defects)
```


# Variables:

# what variables are in the correct data type?

1- Species should be categorical / correct (many wrong imputed values)
2 - Country.of.Origin should be categorical / correct 
3 - Altitude should be integer / incorrect as it is categorical
4 - Number.of.Bags should be integer / incorrect as it is categorical
5 - Bag.Weight should be integer / incorrect as it is categorical
6 - Grading.Date should be integer / incorrect as it is categorical - maybe change the format to only numbers ?
7 - Processing.Method should be factor / incorrect 
8 - Aroma should be double(numerical) / incorrect as it is categorical
9 - Flavor should be double(numerical) / incorrect as it is categorical
10 - Aftertaste should be double(numerical) / incorrect as it is categorical
11 - Acidity should be double(numerical) / incorrect as it is categorical
12 - Body should be double(numerical) / correct
13 - Balance should be double(numerical) / correct
14 - Uniformity should be double(numerical) / correct
15 - Clean.Cup should be double(numerical) / correct
16 - Sweetness should be double(numerical) / correct
17 - Cupper.Points should be double(numerical) / incorrect as it is categorical
18 - Total.Cup.Points should be double(numerical) / incorrect as it is categorical
19 - Moisture should be double(numerical) / incorrect as it is categorical
20 - Category.One.Defects should be factor / incorrect as it is categorical
21 - Quakers should be factor / incorrect as it is categorical - it has one wrong input, maybe delete that 
22 - Color should be categorical / correct 
23 - Category.Two.Defects should be integer / correct - maybe check if this should be factor if it corresponds to how good the quality is 


Perform data type conversion 
- character to factors
- character that hold numbers to double numerical 
- excluding bag weight as the conversion would not work -> both string and "number" and the units are not the same


```{r}
#converting data type as factor
coffee_new_1$Processing.Method<-as.factor(coffee_new_1$Processing.Method)
coffee_new_1$Quakers<-as.factor(coffee_new_1$Quakers)
#converting data type as integer
#coffee_new_1$Altitude<-as.integer(coffee_new_1$Altitude)
#coffee_new_1$Bag.Weight<-as.integer(coffee_new_1$Bag.Weight)

#converting data type as double numerical
coffee_new_1$Number.of.Bags <- as.double(coffee_new_1$Number.of.Bags)

coffee_new_1$Aroma<-as.double(coffee_new_1$Aroma)
coffee_new_1$Flavor<-as.double(coffee_new_1$Flavor)
coffee_new_1$Aftertaste<-as.double(coffee_new_1$Aftertaste)
coffee_new_1$Acidity<-as.double(coffee_new_1$Acidity)
coffee_new_1$Body<-as.double(coffee_new_1$Body)
coffee_new_1$Balance<-as.double(coffee_new_1$Balance)
coffee_new_1$Uniformity<-as.double(coffee_new_1$Uniformity)
coffee_new_1$Clean.Cup<-as.double(coffee_new_1$Clean.Cup)
coffee_new_1$Total.Cup.Points<-as.double(coffee_new_1$Total.Cup.Points)
coffee_new_1$Cupper.Points<-as.double(coffee_new_1$Cupper.Points)
coffee_new_1$Moisture<-as.double(coffee_new_1$Moisture)
coffee_new_1$Category.One.Defects<-as.double(coffee_new_1$Category.One.Defects) 
#checking the structure of the data set with new data types
str(coffee_new_1)
```

validate (before)
```{r}
coffee.rules <- validator(okID = ID>=0, 
                          SpeciesCorrect = is.element(Species,c("Arabica")), 
                          NoNegAltitude = Altitude>0, 
                          NoExtremeAltitude = Altitude< 200000, 
                          NonNegBags = Number.of.Bags>0, 
                          NonNegWeight = Bag.Weight>0, 
                          CorrectMethod = is.element(Processing.Method,
                                                     c("Washed / Wet", "Natural/ Dry", "Pulped natural / honey", "Semi-washed / Semi-pulped", "Other")), 
                          NoNegAroma = Aroma>0, 
                          NoNegFlavor = Flavor>0, 
                          NoNegAftertaste = Aftertaste>0, 
                          NoNegAcidity = Acidity>0, 
                          NoNegBody = Body>0, 
                          NoNegBalance = Balance>0,  
                          NoNegUniformity = Uniformity>0, 
                          NoExtUniformity = Uniformity<=10,  
                          NoNegCup = Clean.Cup>0, 
                          NoExtClean.Cup  = Clean.Cup <=10,  
                          NoNegSweetness = Sweetness>0, 
                          NoExtSweetness = Sweetness<=10, 
                          NoNegPoints = Cupper.Points>0, 
                          NoExtPoints = Cupper.Points<=10, 
                          NoNegTotalPoints = Total.Cup.Points>0, 
                          NoExtTotalPoints = Total.Cup.Points<=100, 
                          NoNegMoisture = Moisture>0, 
                          NoNegQuakers = Quakers>=0, 
                          ColorCorrect = is.element(Color,
                                                    c("Green", "None" , "Bluish-Green", "Blue-Green")),
                          NoMinaltitute = altitude_low_meters>=1,
                          NoextAltitute = altitude_low_meters <= 190164,
                          NoMinaltitute = altitude_high_meters>=1,
                          NoextAltitute = altitude_high_meters <= 190164,
                          NoMinaltitute = altitude_mean_meters>=1,
                          NoextAltitute = altitude_mean_meters <= 190164
) 

out   <- confront(coffee_new, coffee.rules)
out_1 <- confront(coffee, coffee.rules)
summary(out)
plot(out)
```


```{r}
#count NA values
coffee_NA_count <- apply(is.na(coffee_new_1), 2, sum)
coffee_NA_count
# the percentage of NA per variable
coffee_NA_perc <- coffee_NA_count / dim(coffee_new_1)[1] * 100
coffee_NA_perc
```

- Then apply NA imputation to all numerical variables
```{r}
# impute missing values by replacement with the mean value
  # using mutate_if function and if else condition to find any NAs in any numerical columns and replace it with their according median
coffee_new_1 <- coffee_new_1 %>% 
  mutate_if(is.numeric, function(x) ifelse(is.na(x), median(x, na.rm = T), x))
colSums(is.na(coffee_new_1))

median_Body <- median(coffee_new_1$Body, na.rm = T)
coffee_new_1[is.na(coffee_new_1$Body), 7.500] = median_Body
# 
# median_Balance <- median(coffee_new_1$Balance, na.rm = T)
# coffee_new_1[is.na(coffee_new_1$Balance), 7.500] = median_Balance
# 
# median_Sweetness <- median(coffee_new_1$Sweetness, na.rm = T)
# coffee_new_1[is.na(coffee_new_1$Sweetness), 10.000] = median_Sweetness
# 
# median_Uniformity <- median(coffee_new_1$Uniformity, na.rm = T)
# coffee_new_1[is.na(coffee_new_1$Uniformity), 10.00] = median_Uniformity
# 
# median_Clean.Cup <- median(coffee_new_1$Clean.Cup, na.rm = T)
# coffee_new_1[is.na(coffee_new_1$Clean.Cup), 10.000] = median_Clean.Cup

colSums(is.na(coffee_new_1))
```

-dealing with Bag Weight problem
  - all string 
  - Majority is kg put there are some in lbs
  - Created a function to fix this so that all the values are converted to numeric and all have the same unit
```{r}
coffee_new_1
table(coffee_new_1$Bag.Weight)

# replace NA with fake number as the function below is sensitive to NAs 
coffee_new_1$Bag.Weight <- coffee_new_1$Bag.Weight %>% 
  replace_na("99999 kg")

colSums(is.na(coffee_new_1))

# user define function to perform lbs to kg unit conversion with string characters
conversion <- function(x){
  # splitting the string; y is a list
  y <- strsplit(x, " ")
  
  if (y[[1]][2] == "lbs") { #traversing the list within a list
    # Dropping "lbs" units
    x <- gsub(" lbs", "", x)
    # Then performing the unit conversion whilst converting the result to numeric
    x <- 0.45 * as.numeric(x)
  } else {
    # dropping the "kg" unit
    x <- gsub(" kg", "", x)
    # converting the string values to numeric data type
    x <- as.numeric(x)
  }
}
# Applying the function to our bag weight variable
coffee_new_1$Bag.Weight <- sapply(coffee_new_1$Bag.Weight, FUN = conversion)
table(coffee_new_1$Bag.Weight)
str(coffee_new_1)
```

```{r}
# have to change the fake number to NA before doing NA imputation to Bag Weight , will use the naniar function
table(coffee_new_1$Bag.Weight)
library(naniar)
#exp<- coffee_new_1
coffee_new_1 <- coffee_new_1 %>% 
  replace_with_na(replace = list(Bag.Weight = 99999))
table(coffee_new_1$Bag.Weight)
colSums(is.na(coffee_new_1))
```

```{r}
# Then perform NA imputation to Bag weight

median_Bag_weight <- median(coffee_new_1$Bag.Weight, na.rm = T)
coffee_new_1[is.na(coffee_new_1$Bag.Weight), 10.000] = median_Bag_weight


# Then drop any other NA from the character variables
coffee_new_1 <- na.omit(coffee_new_1)
```

- look out for duplicates??
- look out for outliers
- look out for missing data <- imputation
- visualize and show the improvements made

- Maybe here we can show the visual representation of before and after cleaning the data

```{r}
str(coffee_new_1)
table(coffee_new_1$Number.of.Bags)
boxplot(coffee_new_1$Number.of.Bags)
boxplot(coffee_new_1$Bag.Weight)
boxplot(coffee_new_1$Aroma)
boxplot(coffee_new_1$Flavor)
boxplot(coffee_new_1$Aftertaste)
boxplot(coffee_new_1$Acidity)
boxplot(coffee_new_1$Body)
boxplot(coffee_new_1$Balance)
boxplot(coffee_new_1$Uniformity)
boxplot(coffee_new_1$Clean.Cup)
boxplot(coffee_new_1$Sweetness)
boxplot(coffee_new_1$Cupper.Points)
boxplot(coffee_new_1$Total.Cup.Points)
boxplot(coffee_new_1$Moisture)
boxplot(coffee_new_1$Category.One.Defects)
boxplot(coffee_new_1$Quakers)
boxplot(coffee_new_1$Category.Two.Defects)
```

```{r}
table(coffee_new_1$Aroma)
boxplot(coffee_new_1$Aroma)
plot(coffee_new_1$Aroma)
```

```{r}
table(coffee_new_1$Country.of.Origin)
```



# Identify and Remove Outliers

```{r}
str(coffee_new_1)
```


# 1- Country.of.Origin 

```{r}
# function returns the frequency for this variable - makes it possible to identify any outliers
table(coffee_new_1$Country.of.Origin)

#Code to correct the value to Cote d'Ivoire
coffee_new_1["Country.of.Origin"][coffee_new_1["Country.of.Origin"] == "Cote d?Ivoire"] = "Cote d'Ivoire"

#Code to correct the value to United Republic of Tanzania
coffee_new_1["Country.of.Origin"][coffee_new_1["Country.of.Origin"] == "Tanzania, United Republic Of"] = "United Republic of Tanzania"
```

Table function shows that the variable country has one value for "Cote d?Ivoire". This country corresponds to Côte d'Ivoire and it might have been wrongly imputed. Moreover, the same happens with "Tanzania, United Republic Of" that should be United Republic of Tanzania.


# 2- Number.of.Bags 

```{r}
# the variable was not previously converted and it should be numeric type.
coffee_new_1$Number.of.Bags <- as.numeric(coffee_new_1$Number.of.Bags)

# plot to investigate any outliers

ggplot(coffee_new_1) +
  aes(x = Number.of.Bags) +
  geom_histogram(bins = 30L, fill = "#0c4c8a") +
  theme_minimal()

# Use summary function to confirm the outliers - should be the maximum value for this variable 
summary(coffee_new_1$Number.of.Bags)
```

The histogram shows that the variable has one possible outlier -  there seems to be one observation higher than all other observations - it should correspond to the maximum value of this variable. 
To confirm this value, summary function shows a Max of 1062. 
Is then necessary to identify the problem with 1062 number of bags, as it is a value out of the ordinary for this variable.

```{r}
# function to identify which country has 1062 number of bags
subset(coffee_new_1, Number.of.Bags == 1062.0)

# function to understand if it is the norm for Brazil to have 1000 number of bags
subset(coffee_new_1, Country.of.Origin == "Brazil")
```

The output shows that the country Brazil is not set to have 1000 number of bags as the norm. This means that 1062 number of bags is an outlier and needs further investigation.
However, for 1062 number of bags, this variable has 6 quackers, which relates to the poor quality of this set of coffee beans. This explains why there are more number of bags - less quality leads to the need to collect more beans, which leads to more bags.
(It is probably necessary to replace this value by the mean)

```{r}
summary(coffee_new_1$Number.of.Bags)
# maybe check the mean for the number of bags in Brazil?
# Mean = 153.6
```

# 3 - Quakers - unripened coffee beans, often with a wrinkled surface, not darken well when roasted.

```{r}
# Function to get the frequency of this variable - it shows 2 NAs for this variable
coffee_new_1 %>% count(Quakers)

# this function replaces NA by 0
coffee_new_1[is.na(coffee_new_1$Quakers) , ] = 0

# Quakers are not necessarily less than 3, go up and change the validate function. 

ggplot(coffee_new_1, aes(Quakers)) + geom_bar(fill = "#0099F8")

```
The plot shows no outliers.

# 4 - Processing.Method

```{r}
# Function to get the frequency of this variable - it shows 2 missing values for this variable
coffee_new_1 %>% count(Processing.Method)

# output shows NA, this function replaces NA by "Other"
coffee_new_1[is.na(coffee_new_1$Processing.Method) , ] <- "Other"

# function to check for outliers 
ggplot(coffee_new_1, aes(Processing.Method)) + geom_bar(fill="steelblue") + coord_flip()

```

The output shows no outliers for the Processing.Method variable - it also shows that both NA was correctly replaced by "Other".
Moreover, the output also shows that "Washed/Wet" is the most used processing method and that "Pulped natural" is hte least used.


# 5 -  Total.Cup.Points - score coffee on a standardized scale, from 0 to 100.

This variable should range from 0 to 100 and, for a better quality coffee, it should have a value between 80 to 90 (Authentic Coffee Co., 2022)

```{r}
aggregate(coffee_new_1$Total.Cup.Points, list(coffee_new_1$Country.of.Origin), FUN=mean) 
```

The output shows an average of Total.Cup.Points similar to all the countries in this dataset. This means that there are not potential outliers for the Total.Cup.Points variable.

Moreover, the output shows that countries like Cote d'Ivoire, Haiti, India and Nicaragua have a poor Total.Cup.Points value, as it is less than 80.


# 6 - Aroma, Flavor, Aftertaste, Acidity, Body, Balance, Uniformity, Clean.Cup, Sweetness

The coffee quality can be measured by professionals according to several factors such as: Acidity, Aroma, Flavor, Balance, Sweetness, among
others. They also grade each one of these factors in order to get the final score, which indicates the quality of this coffee (Souza, 2022).

Distribution of this variables is presented below.

```{r}
# Plot function to investigate the distribution of each numerical variables and their relationship.
ggplot(coffee_new_1, aes(Aroma, Flavor)) + geom_point()
ggplot(coffee_new_1, aes(Aftertaste, Acidity)) + geom_point()
ggplot(coffee_new_1, aes(Body, Balance)) + geom_point()
ggplot(coffee_new_1, aes(Uniformity, Sweetness)) + geom_point()


# the variable was not previously converted and it should be numeric type.
coffee_new_1$Sweetness <- as.numeric(coffee_new_1$Sweetness)

# the variable was not previously converted and it should be numeric type.
coffee_new_1$Cupper.Points <- as.numeric(coffee_new_1$Cupper.Points)


subset(coffee_new_1, Aroma == 5.08)
subset(coffee_new_1, Acidity == 5.25)
```

The geom_point output above represented show two possible outliers - for variables Aroma and Acidity.
The first one, happens when  Aroma = 5.08 and Acidity = 5.25. However, even thought these two variables are out of the mean values, they stil range between 0-10, which means that they should not be outliers.

(Shows a value of 8.33 for Body when Acidity is 7.08 - even thought is a high value for body when compared with the rest of the variable but it is still within the correct range).

Moreover, the geom_point output does not present any outlier and shows that Flavor-Aroma and Acidity-Aftertaste have a positive correlation between them.


# 7 - Cupper.Points and Clean.Cup

Furthermore, the two variables 
Cupper.Points - Any negative marks that appear in the Uniformity, Clean Cup and Sweetness categories are deducted as two points per cup 
# and
Clean.Cup - refers to a coffee free of taints and defects.

```{r}
ggplot(coffee_new_1, aes(Cupper.Points)) + geom_histogram(binwidth = 0.3)

ggplot(coffee_new_1, aes(Clean.Cup)) + geom_histogram(binwidth = 0.5)

# count function to confirm there are no outliers as represented in the ggplot output
coffee_new_1 %>% count(Cupper.Points)
coffee_new_1 %>% count(Clean.Cup)

table(coffee_new_1$Clean.Cup)
```

The output does not present any possible outlier for these two variables.
Furthermore, as the clean.cup variable corresponds to a coffee free of taints and defects, it makes sense to have 1057 count when clean cup is equal to 10. This means that 1057 coffee beans have no defects.


# 8 - Moisture

```{r}
ggplot(coffee_new_1, aes(Moisture)) + geom_histogram(binwidth = 0.10)

table(coffee_new_1$Moisture)
```

The ggplot output shows that this variable does not present any outlier - most coffee beans have a 0.1 level of moisture.

# 8 - Category.One.Defects and Category.Two.Defects

Bean defects of Category One are rare, as are Quakers.
Bean defects of Category Two are more common and the distribution appears to be approximately log-normal.

```{r}
ggplot(coffee_new_1, aes(Category.One.Defects, Category.Two.Defects)) + geom_boxplot()

ggplot(coffee_new_1, aes(Category.One.Defects)) + geom_histogram(binwidth = 10)
ggplot(coffee_new_1, aes(Category.Two.Defects)) + geom_histogram(binwidth = 10)
```

I do not really understand what these two categories are - could not find anything good on the internet.


# 4. Exploratory Data Analysis

PCA
```{r}
str(coffee_new_1)
colSums(is.na(coffee_new_1))
coffee_pca <- prcomp(coffee_new_1[ ,-c(1,4,18)], center = T, scale = T)
coffee_pca
summary(coffee_pca)
```

```{r}
# calculate the proportion of explained variance (PEV) from the std values
  # sd ^ 2
coffee_pca_var <- coffee_pca$sdev^2
coffee_pca_var
  # PEV is calculated by getting the variance and dividing it by the sum of the variance
coffee_pca_PEV <- coffee_pca_var / sum(coffee_pca_var)
coffee_pca_PEV

plot(coffee_pca)

opar <- par(no.readonly = TRUE)
plot(
  cumsum(coffee_pca_PEV),
  ylim = c(0,1),
  xlab = 'PC',
  ylab = 'cumulative PEV',
  pch = 20,
  col = 'orange'
)
abline(h = 0.8, col = 'red', lty = 'dashed')
par(opar)
```











